---
title: "New Beginning with |>"
description: "Reigniting my data blog using Quarto + GitHub Pages"
author: "Will Hipson"
date: "2023/11/17"
draft: false
categories:
  - R
  - news
  - quarto
  - github
  - serving
---

------------------------------------------------------------------------

New blog, new name, new architecture. Here I talk about my decision to start my blog anew and how I did it.

## A Bit of History

I started my first data [blog](https://willhipson.netlify.app/) back in 2019 when I was a psychology grad student. It was a weird crossroads between academia and programming, but I loved it nonetheless and it offered me a great opportunity to hone my skills and share what I was passionate about.

Shortly after the pandemic hit, I pivoted from academia to data science, landing a data scientist position at a tech start up in the maritime industry. It was my first full time job. I was coding 40 hours a week and enjoying it (most of the time). My hobby became my livelihood and so my blogging frequency declined. A few years later, add a baby to the mix and my blog became a relic.

And on some level, I was fine with this. I'd changed so much in my skills and interests that the blog started to lack cohesiveness. Moreover, I increasingly felt I had less to say in the area of machine learning, but despite having more to say on data engineering and software development, I didn't feel I had sufficient expertise to say anything worthwhile.

## Data Scientist to Data Engineer

Mid 2023 I transitioned from data science to data engineering, landing a new position at the Halifax International Airport Authority. With this has come more freedom to experiment and build community. Hence, the start of this blog.

Why a new blog and not a continuation of the old one? One reason is practical: setting up the development environment for my old blog would likely involve more work than I'm willing to put in. Plus, creating and maintaining a data blog has gotten easier these days with Quarto and GitHub Pages (more on this later).

Another reason is philosophical. I felt it time to make a clean break from my earlier statistical focus. Not that I'm divorcing stats from my work as a data engineer. In fact, by nature of being on a small data team, my work crosscuts almost all aspects of the data landscape, including stats. Still, it's not my main gig these days and so the idea of continuing where I left off just felt a bit kluge-y.

## A Note on the Name

The two characters '\|\>' (individually pronounced pipe + greater than) make up the pipe operator in the R language. `|>` is used for passing an object as the first argument into a function. Using several `|>` to form a *pipeline* is a common design pattern that avoids the need for deeply nested function calls.

So this:

```{r}
# My eyes!!!
nchar(paste(sample(letters, 4), 1:4, collapse = ", "))
```

Becomes this:

```{r}
# Much easier to read
letters |> 
  sample(4) |> 
  paste(1:4, collapse = ", ") |> 
  nchar()
```

To me on a more abstract level, `|>` is like a progression. When I use it for data cleaning I'm often starting with raw, messy data at the start of the pipeline followed by some pipes into `dplyr::filter()` or `dplyr::mutate()`. So I like to think of `|>` as symbolic of my own progression. It also represents my love of the R language - quirks and all.

## Quarto + GitHub Pages

This isn't a tutorial on Quarto or GitHub Pages, rather I want to briefly share my experience moving to this architecture.

In my old blog, I used the `blogdown` package and `rmarkdown` to generate my content, and then I used Netlify for hosting. The initial setup on the Netlify side posed some challenges, particularly to my 2019 coding self. Moving to GitHub Pages seemed a natural choice given that I'd already been deploying my blog via GitHub.

Anyone vaguely familiar with the R ecosystem in recent years will have heard of Quarto. I see it as the next step in the evolutionary chain after RMarkdown. It generalizes a lot of the R-specific stuff that RMarkdown did. Beyond my blog, I've migrated almost all my reporting needs over to Quarto (except for package vignettes and READMEs it seems).

## What to Expect

Where my old blog was strictly an R blog with academic spin, my new blog should be more balanced, still I suspect, with a heavy skew toward R. The focus will be more broadly on the data lifecycle. I think too there will be a bit less polish, with a leaning toward more opinionated pieces instead of how-tos. There are better places for tutorials.

Whether you've been a reader of my old blog or not, I hope you'll follow me in this new blogging venture. It is a joy to write about the vast world of data and software. Even better if you have people who want to listen and share with you.
